---
title: "Extending meow: Writing Your Own Modules"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extending meow: Writing Your Own Modules}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(meow)
```

The whole point of `meow` is that you can drop your own algorithms into a shared
simulation harness and compare them, on equal footing, against existing methods.
This vignette is the authoritative reference for the three kinds of module you
can write: **data loaders**, **item selection functions**, and **parameter
update functions**.

# The simulation state

A `meow` simulation is a single call to `meow()`, which repeatedly alternates
between an item selection step and a parameter update step until no further items
are administered. For speed, the simulation state is represented with **matrices**
rather than long data frames:

| Object  | Type                         | Meaning                                                                 |
|---------|------------------------------|-------------------------------------------------------------------------|
| `R`     | respondents x items matrix   | The (potential) response of every respondent to every item.             |
| `admin` | respondents x items integer  | `0` if an item has not been administered; a positive value if it has.   |
| `pers`  | data frame                   | Person parameter estimates. First column `id`, then one column per parameter. |
| `item`  | data frame                   | Item parameter estimates. First column `item`, then one column per parameter. |
| `adj_mat` | items x items matrix       | Item co-exposure counts (see `construct_adj_mat()`).                     |

Two design choices are worth calling out:

* **Responses are matrices.** Finding the items a respondent has *not* yet seen
  is a single logical comparison (`admin[i, ] == 0`) instead of a join over a
  growing table. The positive values in `admin` additionally encode the **order**
  in which items were administered, which order-sensitive algorithms (such as
  paired-comparison updates) can use.
* **Parameters are data frames.** This preserves the flexibility to carry
  arbitrary person or item parameters: add a column to `pers` or `item` and it
  flows through the simulation untouched.

If you would rather work with long data frames inside your module, call
`meow_long(R, admin)` to get the administered responses as an `id`/`item`/`resp`
data frame, ordered by respondent and administration order.

```{r}
R <- matrix(c(1, 0, 1, 1), nrow = 2)
admin <- matrix(c(1L, 0L, 2L, 1L), nrow = 2)
meow_long(R, admin)
```

# Data loaders

A data loader sets up a simulation. It takes any arguments you like (passed
through `data_args`) and returns a list with three elements:

* `resp` --- a long data frame with columns `id`, `item`, and `resp`.
* `pers_tru` --- true person parameters; first column `id`, then parameters.
* `item_tru` --- true item parameters; first column `item`, then parameters.

`meow()` turns `resp` into the response matrix `R` once, before the loop starts,
so the loader's job is only to produce the bank of potential responses and the
ground-truth parameters.

```{r}
data_two_groups <- function(N_per_group = 50, N_items = 40, data_seed = 1) {
  set.seed(data_seed)
  N <- 2 * N_per_group
  theta <- c(stats::rnorm(N_per_group, -0.5), stats::rnorm(N_per_group, 0.5))
  b <- stats::rnorm(N_items)
  pers_tru <- data.frame(id = seq_len(N), theta = theta)
  item_tru <- data.frame(item = seq_len(N_items), b = b, a = 1)

  p <- stats::plogis(outer(theta, b, "-"))
  resp_mat <- matrix(stats::rbinom(length(p), 1, p), nrow = N)
  resp <- data.frame(
    id = rep(seq_len(N), each = N_items),
    item = rep(seq_len(N_items), times = N),
    resp = as.vector(t(resp_mat))
  )
  set.seed(NULL)
  list(resp = resp, pers_tru = pers_tru, item_tru = item_tru)
}

str(data_two_groups(N_per_group = 3, N_items = 4), max.level = 1)
```

Reserved argument names a loader must **not** use are the ones the harness
supplies to the other modules: `pers`, `item`, `R`, `admin`, `adj_mat`, and
`select_seed`.

# Item selection functions

An item selection function decides which item to administer next. Its signature
is

```r
select_fun <- function(pers, item, R, admin, adj_mat, ...) { ... }
```

and it returns an **administration matrix** with the newly chosen cells marked
non-zero. You only need to *add* items; the harness records the administration
order for you, so setting a cell to `TRUE` or `1` is enough. Leave previously
administered cells as they were. To stop administering items (a stopping rule),
simply return `admin` unchanged --- the simulation halts when an iteration adds
nothing.

Here is a complete custom selector that administers the easiest remaining item to
each respondent (a deliberately naive rule), seeding the first five items on the
first iteration:

```{r}
select_easiest <- function(pers, item, R, admin, adj_mat = NULL) {
  if (!any(admin != 0)) {       # first iteration: seed five items
    admin[, seq_len(min(5, ncol(admin)))] <- 1L
    return(admin)
  }
  difficulty <- item$b
  for (i in which(rowSums(admin == 0) > 0)) {
    remaining <- which(admin[i, ] == 0)
    pick <- remaining[which.min(difficulty[remaining])]
    admin[i, pick] <- 1L
  }
  admin
}
```

A few patterns to notice, all of which avoid touching a growing data frame:

* `admin[i, ] == 0` gives the unadministered items for respondent `i`.
* `rowSums(admin == 0) > 0` identifies respondents who still need an item.
* Setting `admin[i, pick] <- 1L` administers item `pick`.

If you prefer to compute on long data, `meow_long(R, admin)` is always available.

# Parameter update functions

A parameter update function re-estimates parameters from the administered
responses. Its signature is

```r
update_fun <- function(pers, item, R, admin, ...) { ... }
```

and it returns a list with two elements, `pers` and `item`, the updated estimate
data frames. The administered responses are `R[admin != 0]`, and their
respondent/item indices come from `which(admin != 0, arr.ind = TRUE)`.

This custom updater nudges each ability toward a simple proportion-correct
heuristic and leaves items fixed:

```{r}
update_pct_correct <- function(pers, item, R, admin, rate = 0.5) {
  idx <- which(admin != 0, arr.ind = TRUE)
  person <- idx[, 1]
  resp <- R[idx]
  pct <- tapply(resp, person, mean)
  target <- stats::qlogis(pmin(pmax(pct, 0.02), 0.98)) # logit of proportion
  pers$theta[as.integer(names(target))] <-
    (1 - rate) * pers$theta[as.integer(names(target))] + rate * target
  list(pers = pers, item = item)
}
```

# Putting it together

Custom modules plug into `meow()` exactly like the bundled ones. Extra arguments
are passed through `select_args`, `update_args`, and `data_args`.

```{r}
sim <- meow(
  select_fun  = select_easiest,
  update_fun  = update_pct_correct,
  data_loader = data_two_groups,
  data_args   = list(N_per_group = 25, N_items = 20),
  update_args = list(rate = 0.3),
  fix         = "item"
)

head(sim$results[, 1:4])
```

The output object has the same shape no matter which modules you use --- a
`results` data frame (one row per iteration, an `est` and `bias` column for each
parameter), a list of adjacency matrices in `adj_mats`, and the true parameters
in `pers_tru` and `item_tru`. That consistency is what lets you reuse analysis
and plotting code across studies; see `vignette("meow-viz")`.

# Checklist

* Data loaders return `list(resp, pers_tru, item_tru)` with long `resp`.
* Item selection functions take `(pers, item, R, admin, adj_mat, ...)` and return
  an administration matrix with new cells marked non-zero.
* Parameter update functions take `(pers, item, R, admin, ...)` and return
  `list(pers, item)`.
* Use `admin != 0` (or `meow_administered()`) for the administered mask, and
  `meow_long()` if you want long data frames.
* Do not un-administer items, and implement stopping rules by declining to
  administer anything further.
