---
title: "Item Selection Functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Item Selection Functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(meow)
```

Item selection functions determine which item is administered next to each
respondent. They are one of the three swappable components of a `meow`
simulation. For the full module contract, see `vignette("extending-meow")`; this
vignette focuses on the bundled selectors and how to write your own.

# Function signature

Every item selection function has the signature

```r
select_fun <- function(pers, item, R, admin, adj_mat, ...) {
  # ... decide which items to administer ...
  return(admin)
}
```

It receives the current person estimates (`pers`), item estimates (`item`), the
respondent-by-item response matrix (`R`), the integer administration matrix
(`admin`), and the item co-exposure matrix (`adj_mat`). It returns an
administration matrix with the newly chosen cells marked non-zero. The harness
records the order of administration, so you only need to *add* items.

The unadministered items for respondent `i` are `which(admin[i, ] == 0)`, and the
respondents still needing an item are `which(rowSums(admin == 0) > 0)`.

# Bundled selectors

## Sequential

`select_sequential()` administers the lowest-numbered remaining item to each
respondent, producing a fixed linear form.

```{r, eval = FALSE}
select_sequential <- function(pers, item, R, admin, adj_mat = NULL) {
  if (!any(admin != 0)) {
    admin[, seq_len(min(5, ncol(admin)))] <- 1L  # seed the first five items
    return(admin)
  }
  unadmin <- admin == 0
  has <- which(rowSums(unadmin) > 0)
  nextcol <- max.col(unadmin[has, , drop = FALSE] + 0, ties.method = "first")
  admin[cbind(has, nextcol)] <- 1L
  admin
}
```

## Random

`select_random()` draws one remaining item per respondent at random. It accepts a
`select_seed` for reproducibility, which it clears after use.

## Maximum information

`select_max_info()` administers the remaining item with the greatest 2PL Fisher
information, $I(\theta) = a^2 P(\theta)(1 - P(\theta))$, evaluated at each
respondent's current ability estimate. The information for every
respondent-by-item combination is computed as a single matrix, and the maximum is
taken per row.

## Network distance

`select_max_dist()` and `select_max_dist_enhanced()` treat the item pool as a
network whose edge weights are derived from the co-exposure matrix `adj_mat`. They
administer the item farthest (by shortest-path distance) from the items a
respondent has already seen, breaking ties by maximum information. See
`vignette("network-item-selection")`.

# Writing a custom selector

A custom selector need only follow the signature and return an updated `admin`.
Here we administer the item whose difficulty is closest to a respondent's current
ability --- a simple "target the trait" rule.

```{r}
select_targeted <- function(pers, item, R, admin, adj_mat = NULL) {
  if (!any(admin != 0)) {
    admin[, seq_len(min(5, ncol(admin)))] <- 1L
    return(admin)
  }
  for (i in which(rowSums(admin == 0) > 0)) {
    remaining <- which(admin[i, ] == 0)
    gap <- abs(item$b[remaining] - pers$theta[i])
    admin[i, remaining[which.min(gap)]] <- 1L
  }
  admin
}
```

Use it in a simulation by passing it as `select_fun`:

```{r}
sim <- meow(
  select_fun  = select_targeted,
  update_fun  = update_theta_mle,
  data_loader = data_simple_1pl,
  data_args   = list(N_persons = 50, N_items = 30),
  fix         = "item"
)
nrow(sim$results)
```

# Best practices

1. **Handle the first iteration** with `if (!any(admin != 0))` to seed an initial
   set of items.
2. **Never un-administer** an item that has already been given.
3. **Implement stopping rules** by returning `admin` unchanged once a respondent
   should receive no more items.
4. **Prefer matrix operations** over per-row loops where possible; use
   `meow_long()` only when long data is genuinely more convenient.
