---
title: "Parameter Update Functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Parameter Update Functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(meow)
```

Parameter update functions re-estimate person and item parameters from the
responses administered so far. They are the estimation engine of a `meow`
simulation. For the full module contract, see `vignette("extending-meow")`.

# Function signature

Every parameter update function has the signature

```r
update_fun <- function(pers, item, R, admin, ...) {
  # ... re-estimate parameters ...
  list(pers = updated_pers, item = updated_item)
}
```

It receives the current estimates (`pers`, `item`), the response matrix `R`, and
the administration matrix `admin`, and returns a list with the updated `pers` and
`item` data frames. The administered responses are obtained from the matrix
state:

```{r, eval = FALSE}
idx    <- which(admin != 0, arr.ind = TRUE)
person <- idx[, 1]
item_j <- idx[, 2]
resp   <- R[idx]
```

or, equivalently, as a long data frame with `meow_long(R, admin)`.

# Bundled updaters

## Maximum likelihood ability estimation

`update_theta_mle()` treats item parameters as fixed and finds each respondent's
2PL maximum likelihood ability estimate, constrained to $[-4, 4]$. The
log-likelihood is fully vectorized over the administered responses:

```{r, eval = FALSE}
loglik <- function(theta) {
  p <- stats::plogis(item$a[item_j] * (theta[person] - item$b[item_j]))
  sum(resp * log(p) + (1 - resp) * log(1 - p))
}
est <- stats::optim(pers$theta, loglik, lower = -4, upper = 4,
                    method = "L-BFGS-B", control = list(fnscale = -1))
```

## Elo-style updates (Maths Garden)

`update_maths_garden()` updates both abilities and difficulties with the
on-the-fly Elo rule of Klinkenberg, Straatemeier, and van der Maas (2011):

$$\hat\theta_j = \theta_j + K_\theta \sum_i (S_{ij} - E(S_{ij})), \qquad
\hat b_i = b_i + K_b \sum_j (E(S_{ij}) - S_{ij}).$$

See `vignette("maths-garden-update")`.

## Paired Elo updates (Prowise Learn)

`update_prowise_learn()` updates abilities with the same rule, but updates item
difficulties through paired comparisons of consecutively administered items, which
controls rating drift (Vermeiren et al., 2025). See
`vignette("prowise-learn-update")`.

# Writing a custom updater

This updater shrinks each ability estimate toward the mean of the administered
responses (a crude but illustrative rule) and leaves items unchanged.

```{r}
update_shrink <- function(pers, item, R, admin, weight = 0.5) {
  idx <- which(admin != 0, arr.ind = TRUE)
  person <- idx[, 1]
  resp <- R[idx]
  score <- tapply(resp, person, mean)
  target <- stats::qlogis(pmin(pmax(score, 0.02), 0.98))
  who <- as.integer(names(target))
  pers$theta[who] <- (1 - weight) * pers$theta[who] + weight * target
  list(pers = pers, item = item)
}

sim <- meow(
  select_fun  = select_max_info,
  update_fun  = update_shrink,
  data_loader = data_simple_1pl,
  data_args   = list(N_persons = 50, N_items = 30),
  update_args = list(weight = 0.3),
  fix         = "item"
)
nrow(sim$results)
```

# Best practices

1. **Return `list(pers, item)`** --- both data frames, even if one is unchanged.
2. **Bound estimates** to a sensible range to avoid divergence.
3. **Vectorize** over the administered responses (`tapply()`, matrix indexing)
   rather than looping over respondents or items.
4. **Respect administration order** when it matters: `meow_long()` returns
   responses ordered by respondent and then by administration order, which is
   what paired-comparison methods rely on.
